Bad forecast, good decision: what makes a model relevant for determining policy

### 2020-10-27 manuscript/appendix.Rmd
:pen:

## gather models together, indicate true model
sigma_g <- 0.05
models <- list(f1 = f1, f2 = f2, f3 = f3)
model_sigmas <- c(sigma_g, 1.5 * sigma_g, sigma_g)
true_model <- "f3"
```

@@

```{r}
transitions <- lapply(seq_along(models), 
                      function(i) transition_matrices(models[[i]], 
                                                      states, 
                                                      actions, 
                                                      model_sigmas[[i]]))
names(transitions) <- c("f1", "f2", "f3")

@@

## Simulate each policy reps times, with `f3` as the true model:
simulate_policy <- function(i, policy){
  mdp_planning(transitions[[true_model]], reward, discount,
           policy = policy, x0 = x0, Tmax = Tmax) %>%
    select(value, state_index = state, time, action_index = action)  %>% 
    mutate(state = states[state_index])
}

@@

stepahead_unfished <- stepahead_unfished  %>% 
  filter(model != "3") %>% 
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(expected = transitions[[model]][state_index, , 1]  %*% states,
         var = transitions[[model]][state_index, , 1]  %*% states ^ 2 - expected ^ 2,
         low = states[max(which(cumsum(transitions[[model]][state_index,,1]) < 0.025)) ],
         high = states[min(which(cumsum(transitions[[model]][state_index,,1]) > 0.975)) ],
         true = states[next_state ],
         model = paste0("f", model)) 
```

@@

```{r stepahead_fished}
stepahead_fished <- sims %>% 
  filter(model != "3") %>%
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(prob =  transitions[[model]][state_index, next_state, action_index],
         expected = transitions[[model]][state_index, , action_index]  %*% states,
         var = transitions[[model]][state_index, , action_index]  %*% states ^ 2 - expected ^ 2,
         low = states[max(which(cumsum(transitions[[model]][state_index,,action_index]) < 0.025)) ],
         high = states[min(which(cumsum(transitions[[model]][state_index,,action_index]) > 0.975)) ],
         true = states[next_state],
         model = paste0("f", model)) %>%
 select(time, model, true, expected, low, high, var, prob, reps)
```

@@

```{r}
fig1cd <- predictions %>%
  ggplot(aes(x = score, group = model, fill = model)) +
  geom_histogram(binwidth = 2, show.legend = FALSE) +
  coord_cartesian(xlim = c(-100, 1), ylim = c(0,4000)) +
  xlab("Proper score") +
  facet_wrap(~scenario, ncol = 1, labeller =as_labeller(c(
    A_unfished = "C",
    B_fished = "D"))) + 
  scale_x_continuous(breaks = c(-100, 0)) +
  theme(axis.text.y = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
        )

@@

```{r plot_sims}
fig2a <- 
  sims %>%
  filter(time < 25, reps < 5, model != 3) %>%
  ggplot(aes(time, state, col=model, group = interaction(model,reps))) + 
  geom_line(alpha=0.3) 
```

@@

fig2b <- 
npv_df %>%
  filter(model != "3", time %in% seq(1,100, by = 5)) %>%
  ggplot(aes(time, mean_npv)) +
  geom_line(data = optimal, lwd = 1.5, col = "grey20") +
  geom_point(aes(col=model), size = 4, alpha = 0.8) + 
  ylab("Net present value") + xlab("time")

@@

fig3a <- 
  d %>% pivot_longer(names(models), "model") %>%
  ggplot(aes(state, value, col=model)) +
  geom_point() + 
  geom_hline(aes(yintercept = 0)) + 
  coord_cartesian(ylim = c(-5, 8), xlim = c(0,16))
```

@@

```{r plot_policies}
fig3b <- policies %>%
  ggplot(aes(states, escapement, col=model, lty=model)) + 
  geom_line(lwd=2)
```

>>>>

## gather models together, indicate true model
sigma_g <- 0.05
models <- list("1" = f1, "2" = f2, "3" = f3)
model_sigmas <- c(sigma_g, 1.5 * sigma_g, sigma_g)
true_model <- "3"
```

@@

```{r}
transitions <- lapply(seq_along(models), 
                      function(i) transition_matrices(models[[i]], 
                                                      states, 
                                                      actions, 
                                                      model_sigmas[[i]]))
names(transitions) <- c("1", "2", "3")

@@

## Simulate each policy reps times, with `3` as the true model:
simulate_policy <- function(i, policy){
  mdp_planning(transitions[[true_model]], reward, discount,
           policy = policy, x0 = x0, Tmax = Tmax) %>%
    select(value, state_index = state, time, action_index = action)  %>% 
    mutate(state = states[state_index])
}

@@

stepahead_unfished <- stepahead_unfished  %>% 
  filter(model != "3") %>% 
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(expected = transitions[[model]][state_index, , 1]  %*% states,
         var = transitions[[model]][state_index, , 1]  %*% states ^ 2 - expected ^ 2,
         low = states[max(which(cumsum(transitions[[model]][state_index,,1]) < 0.025)) ],
         high = states[min(which(cumsum(transitions[[model]][state_index,,1]) > 0.975)) ],
         true = states[next_state ]) 
```

@@

```{r stepahead_fished}
stepahead_fished <- sims %>% 
  filter(model != "3") %>%
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(prob =  transitions[[model]][state_index, next_state, action_index],
         expected = transitions[[model]][state_index, , action_index]  %*% states,
         var = transitions[[model]][state_index, , action_index]  %*% states ^ 2 - expected ^ 2,
         low = states[max(which(cumsum(transitions[[model]][state_index,,action_index]) < 0.025)) ],
         high = states[min(which(cumsum(transitions[[model]][state_index,,action_index]) > 0.975)) ],
         true = states[next_state]) %>%
 select(time, model, true, expected, low, high, var, prob, reps)
```

@@

```{r}
fig1cd <- predictions %>%
  ggplot(aes(x = score, group = model, fill = model)) +
  geom_histogram(binwidth = 2, show.legend = FALSE) +
  coord_cartesian(xlim = c(-100, 1), ylim = c(0,4000)) +
  xlab("Proper score") +
  facet_wrap(~scenario, ncol = 1, labeller =as_labeller(c(
    A_unfished = "C.",
    B_fished = "D."))) + 
  scale_x_continuous(breaks = c(-100, 0)) +
  theme(axis.text.y = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
        )

@@

```{r plot_sims}
fig2a <- 
  sims %>% 
  group_by(model, time) %>%
  summarise(mean_state = mean(state), sd = sd(state), .groups = "drop") %>%
  filter(time < 25, model != "3") %>%
  ggplot(aes(time, mean_state)) + 
  geom_line(aes(col = model), lwd=1.5, show.legend = FALSE) +
    geom_ribbon(aes(ymin = mean_state - 2*sd, 
                    ymax = mean_state + 2*sd,
                    fill = model),
                alpha = 0.2, show.legend = FALSE) +
  ylab("state") + 
  ggtitle("Ecological performance")
```

@@

fig2b <- 
npv_df %>%
  filter(model != "3", time %in% seq(1,100, by = 5)) %>%
  ggplot(aes(time, mean_npv)) +
  geom_line(data = optimal, lwd = 1.5, col = "grey20") +
  geom_point(aes(col=model), size = 4, alpha = 0.8) + 
  ylab("Net present value") + xlab("time")  + 
  ggtitle("Economic performance")

@@

fig3a <- 
  d %>% pivot_longer(names(models), "model") %>%
  ggplot(aes(state, value, col=model, lty=model)) +
  geom_hline(aes(yintercept = 0), lwd=1) + 
  geom_line(lwd=2, show.legend = FALSE) + 
  coord_cartesian(ylim = c(-5, 8), xlim = c(0,16)) +
  ylab(bquote(f(x))) + xlab("x")
```

@@

```{r plot_policies}
fig3b <- policies %>%
  ggplot(aes(states, escapement, col=model, lty=model)) + 
  geom_line(lwd=2) + xlab("state")
```

### 2020-10-27 manuscript/manuscript.Rmd
:pen:

```{r figure1, fig.width=7, fig.height=5, include=TRUE, message=TRUE, warning=TRUE}
fig1ab + fig1cd + 
  plot_layout(widths =  c(3,1), guides = 'collect') + 
  theme(plot.margin = margin(0, 0, .1, .1, "cm"))
```

@@

```{r figure2, message = FALSE, fig.cap = "Corresponding utility (measured as mean net present value, that is: cumulative value, discounting future values by $\\delta^t$, averaged across replicates). Note that the exepcted utility under model 1, which has the worst forecast, is nearly identical to the optimal utility achieved by managing under the correct model, 3.  The utility derived from model 2 is far smaller, despite it's overall better performance in long term forecasts."}
fig2a + fig2b
```

@@

```{r figure3, fig.width=5, fig.height=7}
fig3a / fig3b
```

>>>>

```{r figure1, fig.width=7, fig.height=5, include=TRUE, message=TRUE, warning=TRUE,, fig.cap = "Forecast performance of each model."}
fig1ab + fig1cd + 
  plot_layout(widths =  c(3,1), guides = 'collect') + 
  theme(plot.margin = margin(0, 0, .1, .1, "cm"))
```

@@

```{r figure2, message = FALSE, fig.cap = "Ecological and economic performance of each forecast."}
fig2a + fig2b
```

@@

```{r figure3, fig.width=7, fig.height=4}
fig3a + fig3b
```
